


---
title: "Suicide Rates"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: spacelab 
---

```{css} 
.value-box { height:100px; 
} 
``` 


```{r setup, include=FALSE}

# Loading the required packages 

library(flexdashboard)
library(knitr) 
library(highcharter) 
library(plotly) 
library(viridis) 
library(tidyverse) 
library(countrycode) 
library(rjson) 
library(crosstalk) 
library(DT) 
# Displaying data tables 
getwd()
setwd("/Users/tanyashetty/Downloads/Data Visualization Assignment")
data <- read.csv('master.csv')  
#Reorder levels of age variable.  
data$age <- factor(data$age, levels = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years")) 

```

```{r include=FALSE} 
# Create a custom theme for the plots.  

custom_theme <- hc_theme( colors = c('#D664BE', 'blue', 'pink'),   chart = list( 
         backgroundColor = '#FAFAFA',           plotBorderColor = "black"),   xAxis = list( 
         gridLineColor = "#070707",  
         labels = list(style = list(color = "#000500")),           lineColor = "#E5E5E5",           minorGridLineColor = "#E5E5E5",           tickColor = "#E5E5E5",  
         title = list(style = list(color = "#000500"))),    yAxis = list(          gridLineColor = "#070707",  
         labels = list(style = list(color = "#000500")),           lineColor = "#E5E5E5",           minorGridLineColor = "#E5E5E5",           tickColor = "#E5E5E5",           tickWidth = 1,  
         title = list(style = list(color = "#000500"))),    
  title = list(style = list(color = '#000500', fontFamily = "Lato")),   subtitle = list(style = list(color = '#666666', fontFamily = "Lato")),   legend = list( 
         itemStyle = list(color = "#000500"),           itemHoverStyle = list(color = "#FFF"),           itemHiddenStyle = list(color = "#606063")),    credits = list(style = list(color = "#666")),   itemHoverStyle = list(color = 'gray')) 
``` 

By World & Continent 
===========================================================

Column {.tabset .tabset-fade data-width=700.colored}
----------------------------------------------------
### Overall Trend {.no-padding}

 
```{r fig.height=5} 
# Code to create tibble for line plot.   
overall_tibble = select(data, year, suicides_no, population) 
overall_tibble = overall_tibble %>%  group_by(year) %>% summarise(suic_capita = round((sum(suicides_no)/sum(population))*100000, 2))   # Code to Create a line plot. 
highchart() %>%  
    hc_add_series(overall_tibble, hcaes(x = year, y = suic_capita, color = suic_capita), type = "line") %>% 
hc_tooltip(crosshairs = TRUE, borderWidth = 1.5, headerFormat = "", pointFormat = paste("Year: {point.x}  Suicides: {point.y}")) %>%     hc_title(text = "Global Trends in Suicide ") %>%      hc_subtitle(text = "1985-2016") %>%     hc_xAxis(title = list(text = "Year")) %>% 
    hc_yAxis(title = list(text = "Suicides per 100K population"),allowDecimals = FALSE) 
``` 


### By Age {.no-padding} 
```{r fig.height=5} 
# Code to create tibble for age variable to use it in the line plot.   
age_tibble <- data %>% 
  select(year, age, suicides_no, population) %>%   group_by(year, age) %>% 
  summarise(suic_capita = round((sum(suicides_no)/sum(population))*100000, 2)) # Code to pick the color for graph.  
age_color <- plasma(6,begin = 0.3, end = 1,direction = -1) # Code to Create a line plot. 
highchart() %>%  
    hc_add_series(age_tibble, hcaes(x = year, y = suic_capita, group = age), type = 
"line", color = age_color) %>% 
    hc_tooltip(crosshairs = TRUE, borderWidth = 1.5, headerFormat = "", pointFormat = paste("Year: {point.x} ","Age: {point.age}", "Suicides: {point.y}")) %>% hc_title(text = "Global Suicide Rates by Age") %>% hc_subtitle(text = "1985-2016") %>% hc_xAxis(title = list(text = "Year")) %>% hc_yAxis(title = list(text = "Suicides per 100K population"), allowDecimals = FALSE) 
``` 

### By Gender {.no-padding} 
```{r fig.height=5.14} 
# Code to create tibble for sex so we can use it when creating our line plot.   
sex_tibble <- data %>% 
  select(year, sex, suicides_no, population) %>%   group_by(year, sex) %>% 
  summarise(suic_capita = round((sum(suicides_no)/sum(population))*100000, 2)) # Code to pick  color for gender variable. 
sex_color <- c("#FFA62B", "#094074") # Orange Peel & Indigo Dye # Code to create line plot. 
highchart() %>%  hc_add_series(sex_tibble, hcaes(x = year, y = suic_capita, group = sex), type = "line", color = sex_color) %>% 
    hc_tooltip(crosshairs = TRUE, borderWidth = 1.5, headerFormat = "", 
pointFormat = paste("Year: {point.x} ","Gender: {point.sex}", "Suicides: {point.y}")) %>% 
    hc_title(text = " Global Suicide Rates by Gender") %>%      hc_subtitle(text = "1985-2016") %>%     hc_xAxis(title = list(text = "Year")) %>% 
    hc_yAxis(title = list(text = "Suicides per 100K population"), allowDecimals = FALSE) 
``` 
           
Row {data-height= 350} 
----------------------------------------------------
### of Suicides per 100K Population <font size='1'> (1985-2016) </font> 
```{r} 
# Code to take the worldwide number of suicides per 100K people from the data 
total_suicides <- "13.12 %" # Create value box 
valueBox(total_suicides, icon = "fa-plus", color = '#D8B0AC') 
``` 

### Suicides by Continent and Gender {.no-title .no-padding .colored } 
```{r fig.height= 4.35} 
# Code to group the Countries based on Continent 
data$continent <- countrycode(sourcevar = data$country,                               origin = "country.name",                               destination = "continent") 
 
 
south_america <- c('Argentina', 'Brazil', 'Chile', 'Colombia', 'Ecuador', 'Guyana', 
'Paraguay', 'Suriname', 'Uruguay') 

data$continent[data$country %in% south_america] <- 'South America'

data$continent[data$continent=='Americas'] <- 'North America' 

# Code to create a tibble for continent and sex. 
continent_sex_tibble <- data %>% 
  select(continent, sex, suicides_no, population) %>%   group_by(continent, sex) %>% 
  summarize(suic_capita = round((sum(suicides_no)/sum(population))*100000, 2)) # Code to create histogram of suicides by continent. 
highchart() %>% 
hc_add_series(continent_sex_tibble, hcaes(x = continent, y = suic_capita, group = sex), type = "column")  %>%      hc_colors(colors = sex_color) %>% 
    hc_title(text = "Suicide Rates by Continent and Gender", style = (list(fontSize = 
'14px'))) %>%  
    hc_subtitle(text = "1985-2016") %>% 
    hc_tooltip(borderWidth = 1.5, pointFormat = paste("Gender:  {point.sex}   Suicides: {point.y}")) %>% 
    hc_xAxis(categories = c("Africa", "Asia", "Europe", "North  America", "Oceania", 
"South  America"), labels = list(style = list(fontSize = 8))) %>%     hc_yAxis(labels = list(style = list(fontSize = 10)),              title = list(text = "Suicides per 100K population",              style = list(fontSize = 10))) 
``` 

### Suicides by Continent and Age {.no-title .no-padding .colored } 

```{r fig.height= 4.35} 
# Code to create a tibble for continent and Age. 
continent_age_tibble <- data %>% 
  select(continent, age, suicides_no, population) %>%   group_by(continent, age) %>% 
  summarize(suic_capita = round((sum(suicides_no)/sum(population))*100000, 2))  # Code to Create histogram of suicides by continent. 
highchart() %>% 
hc_add_series(continent_age_tibble, hcaes(x = continent, y = suic_capita, group = age), type = "column")  %>%      hc_colors(colors = age_color) %>% 
    hc_title(text = "Suicide Rates by Continent and Age", style = (list(fontSize = 
'14px'))) %>%  
    hc_subtitle(text = "1985-2016") %>%     hc_tooltip(borderWidth = 1.5, pointFormat = paste("Age:  {point.age}   Suicides: 
{point.y}")) %>% 
    hc_xAxis(categories = c("Africa", "Asia", "Europe", "North  America", "Oceania", 
"South  America"), labels = list(style = list(fontSize = 8))) %>%     hc_yAxis(labels = list(style = list(fontSize = 10)),              title = list(text = "Suicides per 100K population",                           style = list(fontSize = 10)))  
 
``` 
  


By Country 
===========================================================


Column {data-width=400} 
----------------------------------------------------

### By country {.no-padding} 
```{r fig.height=4.55} 
# Code to create tibble for overall suicides by country 

country_bar <- data %>% 
  select(country, suicides_no, population) %>%   group_by(country) %>% 
  summarise(suic_capita = round((sum(suicides_no)/sum(population))*100000, 2)) %>% arrange(desc(suic_capita)) 

# Code to create an  interactive bar plot 

highchart() %>% hc_add_series(country_bar, hcaes(x = country, y = suic_capita , color = suic_capita ), type = "bar")  %>%      hc_tooltip(borderWidth = 1.5,  
               pointFormat = paste("Suicides: {point.y}")) %>%     hc_legend(enabled = FALSE) %>% 
    hc_title(text = "Suicide Rates by country") %>%      hc_subtitle(text = "1985-2016") %>%     hc_xAxis(categories = country_bar$country,  
             labels = list(step = 1),              min = 0, max = 25, 
             scrollbar = list(enabled = TRUE)) %>% 
    hc_yAxis(title = list(text = "Suicides per 100K population")) %>%     hc_plotOptions(bar = list(stacking = "normal",  
                              pointPadding = 0, groupPadding = 0, borderWidth = 0.5)) %>%      hc_add_theme(custom_theme) 
``` 


References 
=================================================================== 
  
[1]	Kaggle 2018,*Suicide Rates Overview 1985 to 2016*,Compares socioeconomic info with suicide rates by year and country,data file,Kaggle 2018,viewed 22 October 2021, 
<https://www.kaggle.com/russellyates88/suicide-rates-overview-1985-to-2016> *The data was compiled from the four other datasets, the sources are: 
[2]	United Nations Development Program. (2018). Human development index (HDI). Retrieved from <http://hdr.undp.org/en/indicators/137506> 
[3]	World Bank. (2018). World development indicators: GDP (current US$) by country:1985 to 2016. Retrieved from 
<http://databank.worldbank.org/data/source/world-development-indicators#> 
[4]	[Szamil]. (2017). Suicide in the Twenty-First Century [dataset]. Retrieved from 
<https://www.kaggle.com/szamil/suicide-in-the-twenty-first-century/notebook> [5] World Health Organization. (2018). Suicide prevention Retrieved from <http://www.who.int/mental_health/suicide-prevention/en/> 
[6] Baglin,J 2021,'Dashboards',lecture notes,MATH2270,RMIT University,Viewed 
20 October 2021,<https://dark-star-
161610.appspot.com/secured/_book/dashboards.html> 
[7] RDocumentation 2021,*hc_tooltip: Tooltip options for highcharter objects*,viewed 22 October 2021,<https://www.rdocumentation.org/packages/highcharter/versions/0.8.2/to pics/hc_tooltip> 
        
           
           
           



